---
title: "Clickstream analytics @ Data Sci & Eng meetup"
date: "26/5/2018"
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
    toc_float: true
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



# Set-up

* Loading tidyverse mainly to access dplyr for data wrangling in R
* DT a wrapper of the JS library DataTables
* Clickstream is the package that provides most of the functions we need for the Markov Chains analysis



```{r message=FALSE, warning=FALSE, include=TRUE, warnings=F}
library(tidyverse)
library(clickstream)
library(DT)
```



#  Simulate Clickstream data

The analysis was originally run on a high traffic web site. 
In this case we will simulate a fictional ecom website clicksteam


* transition probablility matrix required as input
* additonal parameters include the avg session length and the starting probs
* the result is converted into a data frame for easier inspection


##  Matrix with transition probs

```{r}

my_matrix = matrix(c(0.01, 0.09, 0.05, 0.21, 0.12, 0.17, 0.11, 0.24, 0.00,
                     0.10, 0.00, 0.10, 0.00, 0.00, 0.00, 0.00, 0.80, 0.00,
                     0.07, 0.18, 0.13, 0.05, 0.23, 0.08, 0.26, 0.00, 0.00,
                     0.16, 0.14, 0.07, 0.00, 0.05, 0.22, 0.19, 0.10, 0.07,
                     0.24, 0.17, 0.07, 0.13, 0.00, 0.03, 0.09, 0.26, 0.01,
                     0.11, 0.18, 0.04, 0.15, 0.16, 0.00, 0.10, 0.11, 0.15,
                     0.21, 0.07, 0.08, 0.20, 0.04, 0.18, 0.02, 0.12, 0.08,
                     0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00,
                     0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00), nrow = 9 )

```


## Transition probs data frame

* Remember these probabilities are artificial but we will use clickstream that is generated based on them as if it were created by a natural process




```{r}



transitions <-  t(my_matrix) %>% as.data.frame()

colnames(transitions) <- c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy")

rownames(transitions)<- c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy")


print(transitions)

```

## Generate random clickstreams

* The randomClickstreams function does the work for us. 

* Simulating here a simple ecom website, containing Home page, about page, Product pages etc..

* The simulated data set contains 10K sessions with average length of 15. 

* From there adding a simple sessionID and specified the last step in every session to be either a buy or defer. 

* The result is an object of type clickstream, which is a list. Converting it to a data frame, so that its structured in tabular form. 

```{r }

set.seed(123)



cls <- randomClickstreams(states = c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy"),
                          startProbabilities = c(0.25, 0.10, 0.2, 0.15, 0.1, 0.1, 0.1, 0, 0),
                          transitionMatrix = my_matrix,
                          meanLength = 15, n = 10000)



```





## Clickstream summary

```{r}
summary(cls)
```





## Clickstream in Data Frame form


* optionally add session IDs
* Where no absorbing state is specified, "defer" is added as the last step

we ‘ll get an idea of how the clickstream looks like by printing the data frame 


```{r}

clsdf <- tibble(random_cls=cls) %>% 
    mutate(cls_string = map_chr(random_cls, function(x) paste(unlist(x), collapse = ","))) %>% 
    mutate(cls_complete=ifelse(grepl("Defer|Buy", cls_string), cls_string, paste(cls_string, "Defer", sep=",")))



sessionIDs <- paste0("session_", seq_len(length(cls)))

clsdf$sessionID <- sessionIDs

clsdf %>% dplyr::select(sessionID, cls_complete) %>% head(8)

```


Now lets forget that the data is synthetic and lets treat them as real data. 

Once the data is in this shape we can start to group and count. 



# Frequent paths






## Frequent user paths

By performing a simple group by and count can see which are the most common user paths 
there is one purchase in the top 10 associated with product 3 / we can see the pattern of ppl leaving from the about page. 



```{r}

clsdf %>% rename(user_path=cls_complete) %>% 
  count(user_path) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq)) %>% 
  slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>%  
  datatable()

```


## Frequent paths to conversion

To give more focus on the analysis let's fiter for sessions that end in purchase. 

These leaves us with the most frequent paths to purchase. 

These are paths for which  we want to optimise the site resources, cache strategies, transitions from one page to the other, in order to make them as stable as possible. 


```{r}
clsdf %>% rename(page_sequence=cls_complete) %>% 
  filter(grepl("Buy",page_sequence)) %>% 
  count(page_sequence) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq))  %>% slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  datatable()




```

## Frequent path to non-conversion



```{r}
clsdf %>% rename(page_sequence=cls_complete) %>% 
  filter(grepl("Defer",page_sequence)) %>% 
  count(page_sequence) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq))  %>% slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  datatable()

```





# Clustering 


* This step is optional
* Helps to improve performance of MC model in the next step 


## Prepare object

```{r}
cls_list <- clsdf %>% mutate(cls=map(cls_complete, function(x) unlist(strsplit(x, split = ",")))) 


cls_final <- cls_list$cls


class(cls_final) <- "Clickstreams"


```


## Call k-means algo


```{r}

clusters <- clusterClickstreams(clickstreamList = cls_final, order = 1, centers = 2) 

```



## cluster summary

```{r}
summary(clusters$clusters[[1]]) 
```

## cluster-1 sample


```{r}

sample(clusters[[1]][[1]],size = 5)


```


## cluster-2 sample 

```{r}

sample(clusters[[1]][[2]],size = 5)


```




# Markov Chain Model


* A MC model is applied to fit the data from the second cluster

## fitMarkovChain

```{r}

cluster_two <- clusters$clusters[[2]]

mc <- fitMarkovChain(clickstreamList =cluster_two , order = 1)

summary(mc)

```

## transition probabilities 

```{r}
print(mc)
```








# Visualisation

## Directed Network graph

```{r fig.height=10, fig.width=10}
plot(mc, order = 1, minProbability = 0.09)
```



## Heatmap


```{r}
hmPlot(mc )


```

## Heatmap for absorbing states

```{r}
hmPlot(mc, absorptionProbability = T )
```



# Next click prediction 


```{r}
pattern <- new("Pattern", sequence = c("Prod_3", "Prod_1"),
               absorbingProbabilities = data.frame(Buy = 0.2345288, Defer = 0.7654712))



resultPattern <- predict(mc, startPattern = pattern, dist = 2)

resultPattern


```




