---
title: "Clickstrean Analytics - MeasureCamp #12"
date: "17/3/2018"
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
    toc_float: true
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



# Set-up



```{r message=FALSE, warning=FALSE, include=TRUE, warnings=F}
library(tidyverse)
library(clickstream)
library(arulesSequences)
library(DT)
```



#  Simulate Clickstream data


* transition probablility matrix required as input
* additonal parameters include the avg session length and the starting probs
* the result is converted into a data frame for easier inspection


##  Matrix with transition probs

```{r}

my_matrix = matrix(c(0.01, 0.09, 0.05, 0.21, 0.12, 0.17, 0.11, 0.24, 0.00,
                     0.10, 0.00, 0.10, 0.00, 0.00, 0.00, 0.00, 0.80, 0.00,
                     0.07, 0.18, 0.13, 0.05, 0.23, 0.08, 0.26, 0.00, 0.00,
                     0.16, 0.14, 0.07, 0.00, 0.05, 0.22, 0.19, 0.10, 0.07,
                     0.24, 0.17, 0.07, 0.13, 0.00, 0.03, 0.09, 0.26, 0.01,
                     0.11, 0.18, 0.04, 0.15, 0.16, 0.00, 0.10, 0.11, 0.15,
                     0.21, 0.07, 0.08, 0.20, 0.04, 0.18, 0.02, 0.12, 0.08,
                     0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00,
                     0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00, 0.00), nrow = 9 )

```


## Transition data frame

```{r}



transitions <-  t(my_matrix) %>% as.data.frame()

colnames(transitions) <- c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy")

rownames(transitions)<- c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy")


print(transitions)

```

## Generate random clickstreams

```{r }

set.seed(123)



cls <- randomClickstreams(states = c("Home", "About", "Catalog", "Prod_1", "Prod_2", "Prod_3", "Prod_4", "Defer", "Buy"),
                          startProbabilities = c(0.25, 0.10, 0.2, 0.15, 0.1, 0.1, 0.1, 0, 0),
                          transitionMatrix = my_matrix,
                          meanLength = 15, n = 10000)



```





## Clickstream summary

```{r}
summary(cls)
```





## Clickstream in Data Frame form


* optionally add session IDs
* Where no absorbing state is specified, "defer" is added as the last step

```{r}

clsdf <- tibble(random_cls=cls) %>% 
    mutate(cls_string = map_chr(random_cls, function(x) paste(unlist(x), collapse = ","))) %>% 
    mutate(cls_complete=ifelse(grepl("Defer|Buy", cls_string), cls_string, paste(cls_string, "Defer", sep=",")))



sessionIDs <- paste0("session_", seq_len(length(cls)))

clsdf$sessionID <- sessionIDs

clsdf %>% dplyr::select(sessionID, cls_complete) %>% head(10)

```



# EDA






## Frequent user paths


```{r}

clsdf %>% rename(user_path=cls_complete) %>% 
  count(user_path) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq)) %>% 
  slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>%  
  datatable()

```


## Frequent paths to conversion

```{r}
clsdf %>% rename(page_sequence=cls_complete) %>% 
  filter(grepl("Buy",page_sequence)) %>% 
  count(page_sequence) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq))  %>% slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  datatable()




```

## Frequent path to non-conversion



```{r}
clsdf %>% rename(page_sequence=cls_complete) %>% 
  filter(grepl("Defer",page_sequence)) %>% 
  count(page_sequence) %>% 
  ungroup %>%  
  mutate(freq=n/sum(n)) %>% 
  arrange(desc(freq))  %>% slice(1:20) %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  datatable()

```





# Clustering 


* This step is optional
* Helps to improve performance of MC model in the next step 


## Prepare object

```{r}
cls_list <- clsdf %>% mutate(cls=map(cls_complete, function(x) unlist(strsplit(x, split = ",")))) 


cls_final <- cls_list$cls


class(cls_final) <- "Clickstreams"


```


## Call k-means algo


```{r}

clusters <- clusterClickstreams(clickstreamList = cls_final, order = 1, centers = 2) 

```



## cluster summary

```{r}
summary(clusters$clusters[[1]]) 
```

## cluster-1 sample


```{r}

sample(clusters[[1]][[1]],size = 10)


```


## cluster-2 sample 

```{r}

sample(clusters[[1]][[2]],size = 10)


```




# Markov Chain Model


* A MC model is applied to fit the data from the second cluster

## fitMarkovChain

```{r}

cluster_two <- clusters$clusters[[2]]

mc <- fitMarkovChain(clickstreamList =cluster_two , order = 1)

summary(mc)

```

## transition probabilities 

```{r}
print(mc)
```





# Next click prediction 


```{r}
pattern <- new("Pattern", sequence = c("Prod_3", "Prod_1"),
               absorbingProbabilities = data.frame(Buy = 0.2345288, Defer = 0.7654712))



resultPattern <- predict(mc, startPattern = pattern, dist = 2)

resultPattern


```


# Plots

## Directed Network graph

```{r fig.height=10, fig.width=10}
plot(mc, order = 1, minProbability = 0.09)
```



## Heatmap


```{r}
hmPlot(mc )


```

## Heatmap for absorbing states

```{r}
hmPlot(mc, absorptionProbability = T )
```



# Association rules

##cls to frequencyDF 


```{r}
frequencyDF <- frequencies(cls)

head(frequencyDF, 10)


```

## CLs to transactions

```{r}
trans <- as.transactions(cls)


sequences <- as(cspade(trans, parameter = list(support = 0.3)), "data.frame")


sequences


```



